Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO\textsubscript{2} Reduction Reactions

### 2024-01-09 paper/main_2_abstract.tex
replace math sub/subscript with text

\begin{abstract}
Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
yet the diverse combinations of active elements and support materials, the nature of coordination environments,
elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models to hybrid descriptor based activity volcano plot,
two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
including g-C$_3$N$_4$, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and C$_2$N,
are screened for efficient CO$_2$ reduction reaction (CO$_2$RR).
Starting from establishing a correlation map between the adsorption energies of intermediates and diverse electronic and elementary descriptors, sole singular descriptor lost magic to predict catalytic activity, including d-band center.
Deep learning method utilizing multi-branch CNN model therefore was employed, using 2D electronic density of states (eDOS) as input to predict adsorption energies.
Hybrid-descriptor enveloping both C- and O-types of CO$_2$RR intermediates was introduced to construct volcano plots and limiting potential periodic table,
aiming for intuitive screening of catalyst candidates for efficient CO$_2$ reduction to CH$_4$.
The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
To explore the electronic scale principle governing practical engineering catalytic CO$_2$RR activity,
orbitalwise eDOS shifting experiments based on CNN model were employed.
The study involves examining the adsorption energy and, consequently, catalytic activities while varying supported single atoms.
This work offers a tangible framework to inform both theoretical screening and experimental synthesis,
thereby paving the way for systematically designing efficient SACs.
\end{abstract}

>>>>

\begin{abstract}
Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
yet the diverse combinations of active elements and support materials, the nature of coordination environments,
elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models to hybrid descriptor based activity volcano plot,
two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
including g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N,
are screened for efficient CO\textsubscript{2} reduction reaction (CO\textsubscript{2}RR).
Starting from establishing a correlation map between the adsorption energies of intermediates and diverse electronic and elementary descriptors, sole singular descriptor lost magic to predict catalytic activity, including d-band center.
Deep learning method utilizing multi-branch CNN model therefore was employed, using 2D electronic density of states (eDOS) as input to predict adsorption energies.
Hybrid-descriptor enveloping both C- and O-types of CO\textsubscript{2}RR intermediates was introduced to construct volcano plots and limiting potential periodic table,
aiming for intuitive screening of catalyst candidates for efficient CO\textsubscript{2} reduction to CH\textsubscript{4}.
The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
To explore the electronic scale principle governing practical engineering catalytic CO\textsubscript{2}RR activity,
orbitalwise eDOS shifting experiments based on CNN model were employed.
The study involves examining the adsorption energy and, consequently, catalytic activities while varying supported single atoms.
This work offers a tangible framework to inform both theoretical screening and experimental synthesis,
thereby paving the way for systematically designing efficient SACs.
\end{abstract}

### 2024-01-09 paper/main_3_introduction.tex
major cleanup: flat out dirs as per arxiv request
replace math sub/subscript with text

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{main_sections/figures/main_fig1_pipeline.JPG}
    \caption{\textbf{Catalyst Performance Analysis Pipeline Integrating Volcano Plots and CNN.}
    The figure demonstrates the integrated usage of volcano plots for predictive assessments of existing catalysts and the utility of CNNs to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
    Advanced occlusion and shifting experiments on eDOS are conducted to extract chemical meanings from CNN model.}
    \label{main_fig1:pipeline}
\end{figure}

Electrochemical CO$_2$RR stands as a promising avenue for long-term seasonal energy storage \cite{dinh2018co2} and mitigating excess CO$_2$ in the atmosphere.

Despite decades of research, metallic copper remains a leading CO$_2$RR electrocatalyst for hydrocarbon production \cite{osella2023co2}, albeit burdened by substantial overpotential and low faradic efficiency and poor operational stability \cite{chen2019identifying, liu2021co2}.
In this regard, supported SACs have displayed remarkable activity and selectivity across various catalytic reactions \cite{wang2018heterogeneous, yang2018atomically}, leveraging their unique electronic properties and coordination environment.
To date, experimental studies have primarily focused on two-electron reduction products due to the facile release of gaseous CO \cite{cai2021insights, ju2017understanding, ren2019isolated}.
Consequently, there is a pressing need to modulate the electronic structure of single-atom sites to enable hydrocarbon production at reduced overpotential.

@@

In this study, we introduced a catalyst performance analysis pipeline (\cref{main_fig1:pipeline}) integrating a customized CNN model and volcano plots.
Our CNN model accurately estimated the adsorption energies, the activity descriptor, from the eDOS of 2D materials supported single metal catalysts.
Evaluations on the adsorption energies of nine intermediates along CO$_2$RR and competing hydrogen evolution reaction (HER) as well yielded prediction mean absolute errors (MAEs) at the level of 0.1 eV, which is well aligned with the precision of conventional DFT calculations \cite{kirklin2015open, lejaeghere2016reproducibility, wellendorff2015benchmark}.
This result validated the capability of current CNN model to accurately capture intricate spatial information from eDOS.
Building on this, we introduced an orbitalwise eDOS occlusion experiment was adopted afterwards inspired from the occlusion technique widely used in computer vision and pattern recognition applications.
Systematic eDOS experiment was performed to elucidate orbitalwise contributions to the adsorption energies at different energy levels, which produced consistent outcome from Crystal Orbital Hamilton Population (COHP) analysis and real-space wavefunction visualization.
All these results justified the proficiency of CNN model on identifying the dominant orbitals for adsorbate-substrate interactions.
Furthermore, orbital-wise eDOS shifting experiment, exploring the impact of orbital-wise eDOS shifting on the adsorption behavior of studied catalysts were performed.

@@

In summary, we introduced a catalyst performance analysis pipeline by integrating CNN and volcano plot analysis, using CO$_2$RR as an example proof of concept.
The CNN, which correlates eDOS with activity descriptors, establishes a robust electronic structure-activity relationship.
Notably, it is capable to predict the influence of eDOS disturbances, caused by either alloying, strain, or defects \cite{norskov2011density, bera2017density, kusada2019emergence}, to modulate the catalytic activity of materials, which will offer insights into the intrinsic mechanism for rationalized catalyst optimization strategies.
Advancing the current volcano plot method, the involvement of CNN models in catalyst performance analysis pipeline holds enormous promise for advanced exploration of catalysts with superior performance.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{main_fig1_pipeline.JPG}
    \caption{\textbf{Catalyst Performance Analysis Pipeline Integrating Volcano Plots and CNN.}
    The figure demonstrates the integrated usage of volcano plots for predictive assessments of existing catalysts and the utility of CNNs to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
    Advanced occlusion and shifting experiments on eDOS are conducted to extract chemical meanings from CNN model.}
    \label{main_fig1:pipeline}
\end{figure}

Electrochemical CO\textsubscript{2}RR stands as a promising avenue for long-term seasonal energy storage \cite{dinh2018co2} and mitigating excess CO\textsubscript{2} in the atmosphere.

Despite decades of research, metallic copper remains a leading CO\textsubscript{2}RR electrocatalyst for hydrocarbon production \cite{osella2023co2}, albeit burdened by substantial overpotential and low faradic efficiency and poor operational stability \cite{chen2019identifying, liu2021co2}.
In this regard, supported SACs have displayed remarkable activity and selectivity across various catalytic reactions \cite{wang2018heterogeneous, yang2018atomically}, leveraging their unique electronic properties and coordination environment.
To date, experimental studies have primarily focused on two-electron reduction products due to the facile release of gaseous CO \cite{cai2021insights, ju2017understanding, ren2019isolated}.
Consequently, there is a pressing need to modulate the electronic structure of single-atom sites to enable hydrocarbon production at reduced overpotential.

@@

In this study, we introduced a catalyst performance analysis pipeline (\cref{main_fig1:pipeline}) integrating a customized CNN model and volcano plots.
Our CNN model accurately estimated the adsorption energies, the activity descriptor, from the eDOS of 2D materials supported single metal catalysts.
Evaluations on the adsorption energies of nine intermediates along CO\textsubscript{2}RR and competing hydrogen evolution reaction (HER) as well yielded prediction mean absolute errors (MAEs) at the level of 0.1 eV, which is well aligned with the precision of conventional DFT calculations \cite{kirklin2015open, lejaeghere2016reproducibility, wellendorff2015benchmark}.
This result validated the capability of current CNN model to accurately capture intricate spatial information from eDOS.
Building on this, we introduced an orbitalwise eDOS occlusion experiment was adopted afterwards inspired from the occlusion technique widely used in computer vision and pattern recognition applications.
Systematic eDOS experiment was performed to elucidate orbitalwise contributions to the adsorption energies at different energy levels, which produced consistent outcome from Crystal Orbital Hamilton Population (COHP) analysis and real-space wavefunction visualization.
All these results justified the proficiency of CNN model on identifying the dominant orbitals for adsorbate-substrate interactions.
Furthermore, orbital-wise eDOS shifting experiment, exploring the impact of orbital-wise eDOS shifting on the adsorption behavior of studied catalysts were performed.

@@

In summary, we introduced a catalyst performance analysis pipeline by integrating CNN and volcano plot analysis, using CO\textsubscript{2}RR as an example proof of concept.
The CNN, which correlates eDOS with activity descriptors, establishes a robust electronic structure-activity relationship.
Notably, it is capable to predict the influence of eDOS disturbances, caused by either alloying, strain, or defects \cite{norskov2011density, bera2017density, kusada2019emergence}, to modulate the catalytic activity of materials, which will offer insights into the intrinsic mechanism for rationalized catalyst optimization strategies.
Advancing the current volcano plot method, the involvement of CNN models in catalyst performance analysis pipeline holds enormous promise for advanced exploration of catalysts with superior performance.

### 2024-01-09 paper/main_4.2_volcano_plots.tex
replace math sub/subscript with text

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{main_fig3_volcano.JPG}
    \caption{\textbf{Limiting Potential Volcano Plot and Periodic Table.}
    a. Limiting potential volcano plot computed from hybrid-descriptor-based linear scaling relations at -0.17 V vs SHE.
    In the plot, the circle, triangle, square, diamond, plus, and star symbols represent metals supported on g-$\mathrm{C_3N_4}$, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and $\mathrm{C_2N}$, respectively.
    Dotted purple lines delineate regions representing rate-determining steps (RDS), with roman numerals denoting the respective RDS for each region.
    b. Limiting potential periodic table illustrating the theoretical activity of single atom catalysts, where metals are supported on g-$\mathrm{C_3N_4}$.}
    \label{main_fig3:volcano}
\end{figure}

Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's \cite{peterson2012activity},
goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance \cite{balandin1969modern, deutschmann2000heterogeneous}.
Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor,
which encompasses both C- and O- bonding types of CO$_2$RR intermediates within the scaling relations.
This adaptation arose from the intuition that most CO$_2$RR intermediates are featured with both bonding types rather than any single one in dominance.
By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination ($R^2$) was increased by 0.1387.
Detailed information about the scaling parameters and performance enhancements are provided in Supplementary \cref{si_table13:scaling_params} and Supplementary \cref{si_fig10:r2_hyb_des}.
This improvement to some extent underscored our premise that CO$_2$RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors.
After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately.
Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods,
as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO$_2$RR process, shown in \cref{main_fig3:volcano}a and Supplementary \cref{si_fig11:rds_volcano}.
Given that HER is a key competing reaction with CO$_2$RR \cite{goyal2020competition}, we also incorporated a selectivity volcano plot, presented in Supplementary \cref{si_fig12:sel_volc}, to offer a comprehensive analysis of catalyst candidates.
In adherence to Peterson and Nørskov's original framework \cite{peterson2012activity}, our choice of x and y axes descriptors aligned with $\textit{G}_{\text{ads} \ast \text{CO}}$ and $\textit{G}_{\text{ads} \ast \text{OH}}$.
Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section.
Among the candidatures explored, Ge@g-C$_3$N$_4$ emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV.
Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions.
The RDS volcano plot (Supplementary \cref{si_fig11:rds_volcano}) indicated protonation of *CO as the rate-determining step for nearly optimal candidates.
Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at $\textit{G}_{\text{ads} \ast \text{CO}}$ of 0.0589 eV and $\textit{G}_{\text{ads} \ast \text{OH}}$ of -4.0075 eV, positioned at the right-center of the volcano plot.
Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO$_2$RR.
Consequently, it should be the focal point for future catalyst designs.
The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts.
This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO \cite{peterson2010copper},
and therefore facilitating dynamics and protonation of the *CO species.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.75\linewidth]{main_fig3_volcano.JPG}
    \caption{\textbf{Limiting Potential Volcano Plot and Periodic Table.}
    a. Limiting potential volcano plot computed from hybrid-descriptor-based linear scaling relations at -0.17 V vs SHE.
    In the plot, the circle, triangle, square, diamond, plus, and star symbols represent metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N, respectively.
    Dotted purple lines delineate regions representing rate-determining steps (RDS), with roman numerals denoting the respective RDS for each region.
    b. Limiting potential periodic table illustrating the theoretical activity of single atom catalysts, where metals are supported on g-C\textsubscript{3}N\textsubscript{4}.
    \label{main_fig3:volcano}
\end{figure}

Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's \cite{peterson2012activity},
goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance \cite{balandin1969modern, deutschmann2000heterogeneous}.
Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor,
which encompasses both C- and O- bonding types of CO\textsubscript{2}RR intermediates within the scaling relations.
This adaptation arose from the intuition that most CO\textsubscript{2}RR intermediates are featured with both bonding types rather than any single one in dominance.
By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination ($R^2$) was increased by 0.1387.
Detailed information about the scaling parameters and performance enhancements are provided in Supplementary \cref{si_table13:scaling_params} and Supplementary \cref{si_fig10:r2_hyb_des}.
This improvement to some extent underscored our premise that CO\textsubscript{2}RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors.
After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately.
Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods,
as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO\textsubscript{2}RR process, shown in \cref{main_fig3:volcano}a and Supplementary \cref{si_fig11:rds_volcano}.
Given that HER is a key competing reaction with CO\textsubscript{2}RR \cite{goyal2020competition}, we also incorporated a selectivity volcano plot, presented in Supplementary \cref{si_fig12:sel_volc}, to offer a comprehensive analysis of catalyst candidates.
In adherence to Peterson and Nørskov's original framework \cite{peterson2012activity}, our choice of x and y axes descriptors aligned with $\textit{G}_{\text{ads} \ast \text{CO}}$ and $\textit{G}_{\text{ads} \ast \text{OH}}$.
Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section.
Among the candidatures explored, Ge@g-C\textsubscript{3}N\textsubscript{4} emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV.
Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions.
The RDS volcano plot (Supplementary \cref{si_fig11:rds_volcano}) indicated protonation of *CO as the rate-determining step for nearly optimal candidates.
Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at $\textit{G}_{\text{ads} \ast \text{CO}}$ of 0.0589 eV and $\textit{G}_{\text{ads} \ast \text{OH}}$ of -4.0075 eV, positioned at the right-center of the volcano plot.
Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO\textsubscript{2}RR.
Consequently, it should be the focal point for future catalyst designs.
The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts.
This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO \cite{peterson2010copper},
and therefore facilitating dynamics and protonation of the *CO species.

### 2024-01-09 paper/main_4.3_validation.tex
major cleanup: flat out dirs as per arxiv request
insert si cross-ref in main
replace math sub/subscript with text

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{main_sections/figures/main_fig4_validation.JPG}
    \caption{\textbf{Validation of CNN Predictions.}
    a-c, The occlusion experiment results for (a)Cr $d_{x^2-y^2}$ orbital of Cr@g-$\mathrm{C_3N_4}$, (b)Co $d_{xy}$ orbital of Co@g-$\mathrm{C_3N_4}$ and (c)Cu $d_{xy}$ orbital of Cu@g-$\mathrm{C_3N_4}$.
    Occlusion experiments were conducted on the initial states of CO adsorption process.
    d-f, Real-space wavefunction visualization for (d)Cr $d_{x^2-y^2}$ orbital of Cr@g-$\mathrm{C_3N_4}$, (e)Co $d_{xy}$ orbital of Co@g-$\mathrm{C_3N_4}$ and (f)Cu $d_{xy}$ orbital of Cu@g-$\mathrm{C_3N_4}$.
    Wavefunction visualizations were performed on the final states of CO adsorption process.
    The yellow and aqua colors represent areas of positive and negative Kohn-Sham orbital coefficients, respectively.
    g-j, Occlusion experiment results for (g)$d_{xy}$ orbital and (h)$d_{xz}$ orbital for Co of Co@g-$\mathrm{C_3N_4}$.
    COHP analysis for (i)$d_{xy}$ orbital and (j)$d_{xz}$ orbital for Co of Co@g-$\mathrm{C_3N_4}$, performed between Co and all adsorbate atoms in the final states of CO adsorption process.}
    \label{main_fig4:validation}
\end{figure}

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary Figure 21 (TODO:). Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C$_3$N$_4$, Co@g-C$_3$N$_4$ and Cu@g-C$_3$N$_4$.
Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach.
The results of the occlusion experiment on the supported Cr single atom in Cr@g-C$_3$N$_4$ are given in \cref{main_fig4:validation}a.
The $d_{x^2-y^2}$ orbital emerges as dominant in CO interaction.
Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital.
The real-space wavefunction visualization in \cref{main_fig4:validation}d corroborates these findings, highlighting the interaction between the $d_{x^2-y^2}$ orbital of single metal atom and the adsorbate.
This suggests the $d_{x^2-y^2}$ orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy.
Additionally, occlusion experiments were performed on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$ (\cref{main_fig4:validation}b) and the $Cu d_xy$ orbital of Cu@g-C$_3$N$_4$ (\cref{main_fig4:validation}c).
In both cases, the $d_xy$ orbital was identified as the dominant contributor.
While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well.
Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy \cite{deringer2011crystal}.
In this research, interactions between the supported metal atom and atoms within a 5 \text{\AA} radius were calculated to understand its bonding environment, as detailed in the Methods section.
The results of the occlusion experiment on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in \cref{main_fig4:validation}g and \cref{main_fig4:validation}i.
The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis.
We then investigated the non-dominant Co $d_xz$ orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{main_fig4_validation.JPG}
    \caption{\textbf{Validation of CNN Predictions.}
    a-c, The occlusion experiment results for (a)Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4}, (b)Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and (c)Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Occlusion experiments were conducted on the initial states of CO adsorption process.
    d-f, Real-space wavefunction visualization for (d)Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4}, (e)Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and (f)Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Wavefunction visualizations were performed on the final states of CO adsorption process.
    The yellow and aqua colors represent areas of positive and negative Kohn-Sham orbital coefficients, respectively.
    g-j, Occlusion experiment results for (g)$d_{xy}$ orbital and (h)$d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
    COHP analysis for (i)$d_{xy}$ orbital and (j)$d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and all adsorbate atoms in the final states of CO adsorption process.}
    \label{main_fig4:validation}
\end{figure}

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary \cref{si_fig21:occlusion}. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C\textsubscript{3}N\textsubscript{4}, Co@g-C\textsubscript{3}N\textsubscript{4} and Cu@g-C\textsubscript{3}N\textsubscript{4}.
Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach.
The results of the occlusion experiment on the supported Cr single atom in Cr@g-C\textsubscript{3}N\textsubscript{4} are given in \cref{main_fig4:validation}a.
The $d_{x^2-y^2}$ orbital emerges as dominant in CO interaction.
Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital.
The real-space wavefunction visualization in \cref{main_fig4:validation}d corroborates these findings, highlighting the interaction between the $d_{x^2-y^2}$ orbital of single metal atom and the adsorbate.
This suggests the $d_{x^2-y^2}$ orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy.
Additionally, occlusion experiments were performed on the Co $d_xy$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} (\cref{main_fig4:validation}b) and the $Cu d_xy$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4} (\cref{main_fig4:validation}c).
In both cases, the $d_xy$ orbital was identified as the dominant contributor.
While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well.
Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy \cite{deringer2011crystal}.
In this research, interactions between the supported metal atom and atoms within a 5 \text{\AA} radius were calculated to understand its bonding environment, as detailed in the Methods section.
The results of the occlusion experiment on the Co $d_xy$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4}, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in \cref{main_fig4:validation}g and \cref{main_fig4:validation}i.
The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis.
We then investigated the non-dominant Co $d_xz$ orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

### 2024-01-09 paper/main_5_methods.tex
replace math sub/subscript with text

\subsection{Catalyst models}
Supported SACs were modelled by anchoring individual metal atoms onto various 2D substrates, including graphitic carbon nitride (g-C$_3$N$_4$), nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, boron nitride, and monolayer C$_2$N.
The specific single metal atoms studied are outlined in Supplementary \cref{si_fig1:ptable}.
To prevent interactions between adjacent images along the z-axis, vacuum layers of 15 \text{\AA} were introduced.

\subsection{CO$_2$RR pathway}
The C1 production CO$_2$RR pathway, as initially described by Peterson et al. \cite{peterson2010copper} and later confirmed by Jiao et al. \cite{jiao2017molecular} for g-C$_3$N$_4$ based SACs, involves the following elementary steps.
Asterisks (*) indicate species adsorbed on the SAC surfaces:

@@

The augmented dataset comprises 12,312 samples, with adsorption energies having a standard deviation of 1.8117 eV and a variance of 3.2821 $\mathrm{eV}^2$.
Twenty percent of these samples were allocated to the validation set. Throughout the training process, a batch size of 64 was employed to ensure a balanced learning dynamic.

>>>>

\subsection{Catalyst models}
Supported SACs were modelled by anchoring individual metal atoms onto various 2D substrates, including graphitic carbon nitride (g-C\textsubscript{3}N\textsubscript{4}), nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, boron nitride, and monolayer C\textsubscript{2}N.
The specific single metal atoms studied are outlined in Supplementary \cref{si_fig1:ptable}.
To prevent interactions between adjacent images along the z-axis, vacuum layers of 15 \text{\AA} were introduced.

\subsection{CO$_2$RR pathway}
The C1 production CO$_2$RR pathway, as initially described by Peterson et al. \cite{peterson2010copper} and later confirmed by Jiao et al. \cite{jiao2017molecular} for g-C\textsubscript{3}N\textsubscript{4} based SACs, involves the following elementary steps.
Asterisks (*) indicate species adsorbed on the SAC surfaces:

@@

The augmented dataset comprises 12,312 samples, with adsorption energies having a standard deviation of 1.8117 eV and a variance of 3.2821 eV\textsubscript{2}.
Twenty percent of these samples were allocated to the validation set. Throughout the training process, a batch size of 64 was employed to ensure a balanced learning dynamic.
